---
title: "The Principles of Data Science in Gastroenterology"
---


There are many things to explore in data science for gastroenterology. The data is crucial at all levels and given electronic data has been around for a while in the NHS there is a treasure trove to explore in gastroenterology as well as the other specialities.

*Gastroenterology is event driven

This means that a patient's journey is punctuated by certain events which are recorded as a date time. The combination of a unique identifier and a date time object is fundamental to understanding data manipulation in healthcare


  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

```

##Introduction

The adenoma detection rate (ADR) is a crucial component when assessing the skill of an endoscopist. The rate is attributed to an individual endoscopist and reflects a composite of various techniques that the endoscopist employs such as withdrawl time, use of endocuff etc. However the endoscope used has not been assessed.

##The Problem
<http://rmarkdown.rstudio.com>.


  Data was acquired from retrospectively from hospital records. `r 1+1` (#Filter 1 here) barium swallow reports were analysed. High resolution manometry reports were then acquired and cross referenced with the barium swallow reports. (#Filter 2 here) HRM's with a barium study within 1 year of the barium were retained. Fundoplication and post therapy patients were excluded from the analysis. 
  

   ## Results
    Table 1: Demographics
    
    
    ```{r cars}
    library(xtable)
    knitr::kable(summary(cars))
    ```
    
    ## Including Plots
    
    You can also embed plots, for example:
      
```{r pressure, echo=FALSE}
    plot(pressure)
```
###Code overview (with CodeDepends)
    
 
    
    Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
    

    