---
title: "Grouping by dates"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### How to group

Its fairly common to need to group data by an aspect of the date. For exmaple, how many patients underwent a colonoscopy this month, or week etc.
To do this we need a date object as part of the data as usual. As long as we can extract the aspect we want to group by this should be a breeze with dplyr

###The problem

How do I find out how many endoscopies were done by month for the past calendar year, by endoscopy type

```{r,eval=FALSE}
#Create the data

HospNum_Id<-
Dates<-
EndoscopyType<-
  
df<-as.data.frame(HospNum_Id,Dates,EndoscopyType)
```

###The resulting data looks like this:
```{r}
#library(kable)
#Create the table here

```

Using lubridate, we can extract the month very simply as follows:
This can then be incorporated into dplyr

```{r,eval=FALSE}
#ByMonth<-df %>% select(Dates,EndoscopyType %>% group_by(EndoscopyType,month(Dates)) %>% summarise(n=n())
```

Breaking this down, we are using select() to get the columns we are interested in, then we use group_by() to group according to the two aspects we are interested in
We then use summarise to count each of the groups. The output is then as expected.

More refined analyses may require you to get the first event in a sequence for a patient, or the last. THis again can be done with dplyr as follows:

###Get difference between two dates in consecutive rows
```{r,eval=FALSE}
DateSurveill<-EndoSubsetOnSurveil %>% arrange(HospNum_Id, VisitDate) %>% group_by(HospNum_Id) %>%
  mutate(diffDate = difftime(VisitDate, lag(VisitDate,1)))
```

###Get the first date or the last date in a group
  
