---
title: "More detailed Grouping"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(dplyr)
library(knitr)
```




##Introduction to advanced date groupin


What about finding the difference between dates for two tests for a single patient. We use the same process, but this time the mutate column uses difftime to get the difference between the current row within a group and the previous column (using lag(). If you want the difference between the current and the next row then use lead())
```{r MoreDateGrouping_RowComparison}
Upstage<-df %>% arrange(HospNum_Id, dat) %>% group_by(HospNum_Id) %>%
  mutate(diffDate = difftime(dat, lag(dat,1),units = "days"))
kable(head(Upstage,25))
  
``` 

What if you simply want to plot out the number of procedures done by year. That needs you to extract the year from the date and then summarise as follows:

```{r MoreDateGrouping_NumProcPlot}
Tots<-df %>%
  mutate(year = format(dat, "%Y")) %>%
  group_by(year)%>%
  summarise(n = n())

kable(Tots)
``` 

of course you can mix in other things into dplyr, and this could have been done using the in-built r function table()

```{r MoreDateGrouping_NumberProcTable}
Tots<-df %>%
  mutate(year = format(dat, "%Y"))
  mytable<-table(Tots$year)
  mytable
``` 