---
title: "Making Data from Data with dplyr::Mutate"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                       warning=FALSE, message=FALSE)
library(knitr)
library(kableExtra)
library(dplyr)
```
There are many occasions when a column of data needs to be created from an already existing column for ease of data manipulation. For example, perhaps you have a body of text as a pathology report and you want to extract all the reports where the diagnosis is dysplasia. You could just subset the data using grepl so that you only get the reports that mention this word...but what if the data needs to be cleaned prior to subsetting like excluding reports where the diagnosis is normal but the phrase 'No evidence of dysplasia' is present. Or perhaps there are other manipulations needed prior to subsetting.

This is where data accordionisation is useful. This simply means the creation of data from (usually) a column  into another column in the same dataframe.

The neatest way to do this is with the mutate function from the 'dplyr' package which is devoted to this. There are also other ways which I will demonstrate at the end.


<br><br>
 
**The input data here will be an endoscopy data set:**


```{r DataFromData_Input}
Age<-sample(1:100, 130, replace=TRUE)
Dx<-sample(c("NDBE","LGD","HGD","IMC"), 130, replace = TRUE)
TimeOfEndoscopy<-sample(1:60, 130, replace=TRUE)


EMRdf<-data.frame(Age,Dx,TimeOfEndoscopy,stringsAsFactors=F)
```


Perhaps you need to calculate the number of hours spent doing each endoscopy rather than the number of minutes

```{r DataFromData_SimpleMutate}
EMRdftbb<-EMRdf%>%mutate(TimeOfEndoscopy/60)
#Just show the top 20 results
kable(head(EMRdftbb,20))
```

That is useful but what if you want to classify the amount of time spent doing each endoscopy as follows: <0.4 hours is too little time and >0.4 hours is too long.

<br><br>
 
##Using ifelse() with mutate for conditional accordionisation


 
For this we would use ifelse(). However this can be combined with mutate() so that the result gets put in another column as follows

```{r DataFromData_ifelseMutate}
EMRdf2<-EMRdf%>%mutate(TimeInHours=TimeOfEndoscopy/60)%>%mutate(TimeClassification = ifelse(TimeInHours>0.4, "Too Long", "Too Short"))
#Just show the top 20 results
kable(head(EMRdf2,20))
```

Note how we can chain the mutate() function together.

<br><br>
 
##Using multiple ifelse() 
What if we want to get more complex and put several classifiers in? We just use more ifelse's:


```{r DataFromData_ifelseMultiple}
EMRdf2<-EMRdf%>%mutate(TimeInHours=TimeOfEndoscopy/60)%>%mutate(TimeClassification = ifelse(TimeInHours>0.8, "Too Long", ifelse(TimeInHours<0.5,"Too Short",ifelse(TimeInHours>=0.5&TimeInHours<=0.8,"Just Right","N"))))
#Just show the top 20 results
kable(head(EMRdf2,20))
```

<br><br>
                
##Using multiple ifelse() with grepl() or string_extract

Of course we need to extract information from text as well as numeric data. We can do this using grepl or string_extract from the library(stringr). We have used this before [here](http://gastrodatascience.com/Text_Clean.html) so you may want to refresh yourself.

Let's say we want to extract all the samples that had IMC. We don't want to subset the data, just extract IMC into a column that says IMC and the rest say 'Non-IMC

Using the dataset above:
```{r DataFromData_ifelseGrepl}
library(stringr)
EMRdf$MyIMC_Column<-str_extract(EMRdf$Dx,"IMC")

#to fill the NA's we would do:
EMRdf$MyIMC_Column<-ifelse(grepl("IMC",EMRdf$Dx),"IMC","NoIMC")
#Another way to do this (really should be for more complex examples when you want to extract the entire contents of the cell that has the match)

EMRdf$MyIMC_Column<-ifelse(grepl("IMC",EMRdf$Dx),str_extract(EMRdf$Dx,"IMC"),"NoIMC")
```


So data can be usefully created from data for further analysis


